### 2011-08-30 Text/paper.tex
More explanations and definitions of the mass transfer.

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amssymb,amsmath}
\usepackage{bm}
\usepackage{color}
\usepackage{tabularx}
\begin{document}
\section{Motivation}
The mass transfer is a very important characteristic for microchannels  design. It allows to
properly manufacture a microchannel with necessary properties to ensure that chemical
reactions are performed in the best manner. In
what follows the mass transfer associated with the bubble moving in the microchannel through liquid
medium is studied.

@@

>>>>

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amssymb,amsmath}
\usepackage{bm}
\usepackage{color}
\usepackage{tabularx}
\newcommand{\beq}{\begin{equation}}
\newcommand{\feq}{\end{equation}}
\newcommand{\beqstar}{\begin{equation*}}
\newcommand{\feqstar}{\end{equation*}}                                     
\newcommand{\beqal}{\begin{equation}\begin{aligned}}
\newcommand{\feqal}{\end{aligned}\end{equation}}
\newcommand{\beqalstar}{\begin{equation*}\begin{aligned}}
\newcommand{\feqalstar}{\end{aligned}\end{equation*}}
\begin{document}
\section{Motivation}
The mass transfer is a very important characteristic for microchannels  design. It allows to
properly manufacture a microchannel with necessary properties to ensure that chemical
reactions are performed in the best manner. In
what follows the mass transfer associated with the bubble moving in the microchannel through liquid
medium is studied.

@@

\section{Mass transfer definitions}
By the definition the mass transfer coefficient is the following:
\beq
k_L=\frac{\dot{m}}{P \Delta C},
\feq
where $\dot{m}$ is the mass rate $\Bigl[\frac{kg}{s}\Bigr]$, $P$ is the area of the surface
$\Bigl[m^2\Bigr]$, $\Delta C$ is the concentration difference $\Bigl[\frac{kg}{m^3}\Bigr]$.
Therefore, $k_L$ has a dimension of velocity $\Bigl[\frac{m}{s}\Bigr]$.

There are different methods to estimate the mass transfer coefficient $k_L$. We first examine the
theoretical definitions of the mass transfer in case of the point sources.
\subsection{Point mass sources}
Let us look at the infinitisemal small domain of the volume $A \Delta x$ in a steady-state, not
moving and with the mass source. Then the concentration difference can be found as $\Delta C = C^* -
C(t)$, where $C*$ is the imposed concentration on the infinitesimal concentration source, $C(t)$ is
the time depending average concentration. Therefore one can write the time dependent ODE for the
average concentration in the domain:
\beq
\dot{m}= A \Delta x \frac{\mathrm{d}C}{\mathrm{d} t} = k_L P (C^{*}-C(t)), 
\feq
with the initial condition $C(0)=0$
The solution in time becomes as follows:
\beq
C(t)= C^{*}(1-\exp(-k_L a t )), 
\feq
where $k_L a$ is the volumetric mass transfer coefficient.

This simple picture is able to predict mass transfer happening in moving with the velocity $U$
liquid, see Fig. \ref{fig:moving:frame}.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/mass_transfer.eps}
\caption{The mass transfer in the moving liquid. \label{fig:moving:frame}}
\end{figure}

The mass accumulated in the unit volume $V=A \Delta x$ can be calculated as the difference of 
mass fluxes entering and leaving domain $U \bigl(C(x+\Delta x)-C(x)\bigr)$. We assume that the
source is a point source with the effective perimeter P. The accumulated mass should be proportional
to the mass transfer coefficient:
\beq
U \bigl(C(x+\Delta x)-C(x)\bigr)=k_L P (C^{*}-C(x)), 
\feq 
giving the same equation but only in the coordinate domain:
\beq
C(x)= C^{*}(1-\exp(-k_L a \frac{x}{U} )).
\label{main:mass:transfer:expression} 
\feq
The expression (\ref{main:mass:transfer:expression}) is exactly used in the experiments by
\citet{bercic-mass}. One should be accurate with the definition of velocities because we have two
different phases. Usually, if we cannot distinguish phases one can take the velocity $U$ to be as
average velocity or $U=U_{sl}+U_{sg}$, where $U_{sl}$ and $U_{sg}$ are liquid and gas superficial
velocities correspondingly. However, the real picture is much more complicated, as two phases can
have different mixing patterns from film to slug and from slug to film. For example, one of the
most important characteristics to properly analytically estimate the mass transfer coefficient is
the mixing patterns. For example, one can assume that the slug concentration is uniformly
distributed over the slug, which usually overpredicts the mass transfer \cite{irandoust} since some
tracer concentration from film is mixed with the slug. Another issue is that if the slug is long
enough then the film saturates with the tracer concentration and its influence on the mass transfer
can be negligible \cite{vanbaten-circular}. Mixing patterns of the film and the liquid slugs are of
great importance for the analytical estimation of the mass transfer \cite{yue-mass}.

\subsection{Numerical calculation}
In the numerical simulation it is not possible to simulate really long domains with the infinite
number of the bubbles. Therefore, one needs to come up with the simple smaller domain calculations
of the mass transfer coefficient. The mixing patterns are extremely important for the analytical
estimations of the mass transfer coefficient. This is not the case for numerical calculations,
where one needs to present the adequate picture and the way to calculate the volumetric mass
transfer coefficient. We consider the following possibilities of conducting numerical simulations:
\begin{enumerate}
 \item The tracer picking up procedure can be introduced in the manner of the considerations above.
We impose the concentration at the entrance to be constant $C(0,y)=C_{inlet}$, at the outlet we
impose the outflow boundary condition $\frac{\partial C}{\partial x}(L,y)=0$. One can se the sketch
of the approach in Fig. \ref{fig:benchmark:jos}. Without loss of the generality one can impose
$C_{inlet}=0$. After that one can estimate the mass transfer coefficient through the mass flux
depending on the coordinate $x$. The mass flux is calculated through the following equation:
\beq
j(x)=\int{U(x) C(x,y) \mathrm{d}y}.
\feq
As it was indicated the mass flux is proportional to the mass transfer coefficient $j(x)=k_L \Delta
C$, where the difference in the concentration is taken as $\Delta C=C^{*}-C_{media}$, where $C^{*}$
is the saturation concentration. The question remains with the $C_{media}$ which can be treated
differently. According to the Fick's law $C_{media}$ should be the concentration of the media
examined. However, that imposes the assumption of the ideal mixing in the liquid slug. Another
possibility to take $C_{media}$ as the outflow concentration. A few considerations:
\begin{itemize}
 \item Note, that in the presence of the
complicated geometries and non-homogenous concentrations distributions around the bubble one can
use any definitions. Though the definition of the mass averaged $C_{media}$ is prefered.  
\item Note, that if the tracer volume is large then $C_{media}$ can be close to zero.
\item Note, that one cannnot use the definition of the mass transfer coefficient as far as it
varies through the bubble due to inhomogenoities of the flow. Therefore, the volumetric mass
transfer coefficient is introduced that takes into the account the bubble geometry.
\end{itemize}

\begin{figure}
\includegraphics[width=\textwidth]{Figures/benchmark_jos.eps}
\caption{The numerical mass transfer benchmark. \label{fig:benchmark:jos}}
\end{figure}
\item The periodic domain.
\end{enumerate}
